---
title: "Project 4"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
    theme: 
      version: 4
      bootswatch: litera
runtime: shiny
---

```{r global}
library(here)
library(flexdashboard)
library(DT)
library(tidyverse)
library(maps)

# check if a directory named data exists locally
if (!dir.exists(here("data"))) {
  # if not, create it programmatically
  dir.create(here("data"))
}

# check whether we have the input data
if (!file.exists(here("data", "salary_potential.rda"))) {
    # If we don't, then download the data
    salary_potential <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv")

    save(salary_potential, file = here("data", "salary_potential.rda"))
}

# load the data
load(here("data", "salary_potential.rda"))
```

About
=====================================

## Column 1 {data-width=400}

**Purpose**

In this dashboard, I will analyze a dataset about estimated career pay for college students in the US. The goal here is to address the following questions:

* How does the early career pay differ across the U.S.? 
* What is the relationship between early career pay and mid career pay? 
* What is the relationship between early career pay and STEM percent of student body?

I obtained the original dataset from  [TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10). I referred to this [data dictionary](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10#data-dictionary) from TidyTuesday.

## Column 2 {data-width=600}

[insert Youtube tutorial]

The Data
=====================================

## Column 1 {.sidebar}

Description of the Data

## Column 2 {ata-width=650}

### Table 1

```{r}
salary_potential %>%
  DT::datatable()
```

Table {data-navmenu="Pay by State"}
=====================================

```{r}
# calculate the average estimated early career pay for each state
salary_st_avg <- salary_potential %>%
  select(name, state_name, early_career_pay) %>%
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  group_by(region) %>%
  summarise(
    avg_early_career_pay = mean(early_career_pay)
  )
  
# find coordinates for us states
us_map <- map_data("state") %>%
  mutate(region = gsub(" ", "-", region))

# add salary info to the map
map_salary_st_avg <- us_map %>%
  left_join(salary_st_avg, by = join_by(region))
```

## Column 1 {.sidebar}

Description of the Data

## Column 2 

### Table 1

```{r}
salary_st_avg %>%
  DT::datatable()
```

Plot {data-navmenu="Pay by State"}
=====================================

## Column 1 {.sidebar}

Description of the Data

## Column 2 

### Chart 1

```{r}
# plot the average estimated early career pay across the US
map_salary_st_avg %>% ggplot() +
  geom_polygon(aes(x = long, y = lat, group = group, fill = avg_early_career_pay)) +
  scale_fill_gradient(low = "#FAAB18", high = "#1380A1") +  # Adjust the color scale
  labs(title = "The Average Estimated Early Career Salary Across the U.S.", subtitle = str_wrap("California, Texas, and the states in the northeastern region have higher estimated early career pay compared to states in the central portion of the U.S.", 60), caption = "Yuhan Xiao", fill = "Salary") +
  theme_void(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        plot.subtitle = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"))
```


Trend
=====================================

```{r}
# select salary info from states of interest
salary <- salary_potential %>%
  select(state_name, early_career_pay, mid_career_pay) %>%
  rename(state = state_name) %>%
  mutate(state = tolower(state)) 

# break the data set down into groups defined by states
by_state <- split(salary, salary$state)

# calculate average early pay and mid pay for each state
early_pay <- map_dbl(by_state, ~ mean(.x$early_career_pay, na.rm=TRUE))
mid_pay <- map_dbl(by_state, ~ mean(.x$mid_career_pay, na.rm=TRUE))

# combine average pay into a table
pay_trend <- tibble(
  state = names(early_pay),
  avg_early_pay = early_pay,
  avg_mid_pay = mid_pay
)
```

## Column 1 {.sidebar}

Description of the Data

## Column 2 

### Chart 1

```{r}
# plot the association between average early career pay to average mid career pay
pay_trend %>% 
  ggplot() +
  geom_point(aes(x = avg_early_pay, y = avg_mid_pay)) +
  labs(title = str_wrap("The Association Between Average Estimated Early Career Salary and Mid Career Salary by State", 60), subtitle = str_wrap("A higher early career salary is associated with a higher mid career salary", 60), x = "average early career salary", y = "average mid career salary", caption = "Yuhan Xiao") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 0, hjust = 1),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"),
        legend.position = "none")
```


STEM - Early Pay
=====================================

```{r}
# select salary info from states of interest
pay_stem_trend <- salary_potential %>%
  rename(state = state_name) %>%
  mutate(state = tolower(state)) 

state_choices <- pay_stem_trend %>%
  select(state) %>%
  distinct(state)
```

## Column 1 {.sidebar}

Description of the Data

```{r}
selectInput("state_choice", "Select a state:",
            state_choices)
```

## Column 2 

### Chart 1

```{r warning = F, message = F, fig.dim = c(8,6)}
renderPlot({
  pay_stem_trend %>% 
  filter(state == input$state_choice) %>%
  ggplot(aes(x=stem_percent, y = as.integer(early_career_pay))) +
  geom_point() +
  geom_smooth() + 
  facet_wrap(~state, nrow = 2) +
  labs(title = str_wrap("The Association Between Average Estimated Early Career Salary and Percent of STEM students by State", 60), subtitle = str_wrap("A higher percent of student body in STEM is associated with a higher estimated early career salary", 60), x = "STEM percent", y = "average early career salary", caption = "Yuhan Xiao") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 0, hjust = 1),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"),
        legend.position = "none")
})
```

STEM - Mid Pay
=====================================

## Column 1 {.sidebar}

Description of the Data

```{r}
selectInput("state_choice", "Select a state:",
            state_choices)
```

## Column 2 

### Chart 1

```{r warning = F, message = F, fig.dim = c(8,6)}
renderPlot({
  pay_stem_trend %>% 
  filter(state == input$state_choice) %>%
  ggplot(aes(x=stem_percent, y = as.integer(mid_career_pay))) +
  geom_point() +
  geom_smooth() + 
  facet_wrap(~state, nrow = 2) +
  labs(title = str_wrap("The Association Between Average Estimated Mid Career Salary and Percent of STEM students by State", 60), subtitle = str_wrap("A higher percent of student body in STEM is associated with a higher estimated mid career salary", 60), x = "STEM percent", y = "average mid career salary", caption = "Yuhan Xiao") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 0, hjust = 1),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"),
        legend.position = "none")
})
```

Analysis {style="position:relative;"}
=====================================

## Column 1 {.sidebar}

**Summary**

Colleges have higher estimated early career salary for students if they are located in California, Texas, the northeastern region or if they have a higher percent of student body in STEM. A higher early career salary is associated with a higher mid career salary.

\* Chart 1 Note: 48 states and District of Columbia (D.C.) are included in this map. Alaska and Hawaii are not included in this map.

## Column 2 {style="height:100pc;"}

### Chart 1 *

```{r}
# plot the average estimated early career pay across the US
map_salary_st_avg %>% ggplot() +
  geom_polygon(aes(x = long, y = lat, group = group, fill = avg_early_career_pay)) +
  scale_fill_gradient(low = "#FAAB18", high = "#1380A1") +  # Adjust the color scale
  labs(title = "The Average Estimated Early Career Salary Across the U.S.", subtitle = str_wrap("California, Texas, and the states in the northeastern region have higher estimated early career pay compared to states in the central portion of the U.S.", 60), caption = "Yuhan Xiao", fill = "Salary") +
  theme_void(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        plot.subtitle = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"))
```

### Chart 2

```{r}
# plot the association between average early career pay to average mid career pay
pay_trend %>% 
  ggplot() +
  geom_point(aes(x = avg_early_pay, y = avg_mid_pay)) +
  labs(title = str_wrap("The Association Between Average Estimated Early Career Salary and Mid Career Salary by State", 30), subtitle = str_wrap("A higher early career salary is associated with a higher mid career salary", 40), x = "average early career salary", y = "average mid career salary", caption = "Yuhan Xiao") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 0, hjust = 1),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"),
        legend.position = "none")
```

### Chart 3

```{r warning = F, message = F, fig.dim = c(8,6)}
pay_stem_trend %>% 
  ggplot(aes(x=stem_percent, y = early_career_pay, color = state)) +
  geom_point() +
  geom_smooth() + 
  facet_wrap(~state, nrow = 2) +
  labs(title = str_wrap("The Association Between Average Estimated Early Career Salary and Percent of STEM students by State", 60), subtitle = str_wrap("A higher percent of student body in STEM is associated with a higher estimated early career salary", 60), x = "STEM percent", y = "average early career salary", caption = "Yuhan Xiao") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 0, hjust = 1),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"),
        legend.position = "none")
```

